---
title: "Homework 2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```


### Problem 0
 
Created an R project - p8105_hw2_rp3022 and a subdirectory "data" to store all the files to be used for the homework.
Loading the required libraries

```{r load_libraries}
library(tidyverse)
library(readxl)
```

## Problem 1


```{r}
trans_ent = 
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```
```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
### Answers for Q1
* There are 465 distinct stations.
* There are 84 stations that are ADA compliant
* 37.7% of station entrances and exits without vending allow entrance.
* 60 distinct stations serve on the A train. Of the stations that serve A train, 17 are ADA compliant


## Question 2

```{r Mr. trashwheel}
trash_wheel_mr =read_excel("data/Trash Wheel Collection Data.xlsx", sheet= "Mr. Trash Wheel", range ="A2:N549")%>%
  janitor:: clean_names()%>%
  drop_na(dumpster)%>%
  mutate (sports_balls = as.integer(sports_balls), wheel="mr")

```
```{r Professor trashwheel}
trash_wheel_prof = read_excel("data/Trash Wheel Collection Data.xlsx", sheet= "Professor Trash Wheel", range ="A2:M96") %>%
  janitor::clean_names() %>% 
  drop_na(dumpster)%>%
  mutate(year=as.character(year), wheel="professor")
```

```{r combine}
combined_trash_wheel = bind_rows(trash_wheel_mr, trash_wheel_prof)%>%
  janitor::clean_names()
```


`r skimr::skim(combined_trash_wheel)`

### Description of the dataset:

The total number of observations (rows) are __`r nrow(combined_trash_wheel)`__ and the total number of variables(columns) are **`r ncol(combined_trash_wheel)`**.The names of the variables are **`r colnames(combined_trash_wheel)`** The **numeric variables** are 
weight_tons : the weight of trash collected, 
volume_cubic_yards : volume of trash collected in cubic yards,
and the variables to calculate number of plastic bottles , polystyrene, cigarette butts, glass bottles, grocery bags, chips bags, sports balls collected and number of homes empowered.
The **character variables** are year and month and there is a **date variable**.
The total weight of trash collected by Professor trash wheel is **`r sum(trash_wheel_prof$weight_tons, na.rm = TRUE)`** tons
The total number of sports balls collected by Mr. Trash Wheel is **`r sum(trash_wheel_mr[which(trash_wheel_mr$year == "2020" ), "sports_balls"])`**

## problem 3

```{r polsmonth}
pols_df = read_csv("data/fivethirtyeight_datasets/pols-month.csv")%>%
  janitor::clean_names()%>%
   separate(col=mon, into=c(  "year", "month", "day"), sep = '-', convert = TRUE)%>%
  mutate(month = month.abb[month],
         president = case_when (prez_gop== 1 ~ "gop", prez_dem== 1 ~"dem"))%>%
  select(-prez_gop, -prez_dem, -day)
```

```{r}
snp_df = read_csv("data/fivethirtyeight_datasets/snp.csv")%>%
  janitor::clean_names()%>%
  separate(col=date, into=c("month", "day", "year"), sep ='/', convert=TRUE)%>%
  mutate(month = month.abb[month],
          year = ifelse(year > 49, year+1900, year+2000))%>%
  select(-day)%>%
  select(year, month, everything())
```
```{r}
unemp_df = read_csv(
  "data/fivethirtyeight_datasets/unemployment.csv")%>%
  janitor:: clean_names()%>%
  pivot_longer(jan:dec,
               names_to = "month",
               values_to = "unemployment")%>%
  mutate(month = str_to_title(month))
```
```{r}
pomo_snp = left_join(pols_df, snp_df)
```
```{r}
pomo_unemp = left_join(pomo_snp, unemp_df)
```

* The dataset **pols-month.csv** describes the number of national politicians who were either democratic or republic at a given time. It initially had 822 observations and 9 variables. Individual columns of "prez_dem" and "prez_gop" were combined to create "president" column.  The variable mon was divided into 3 variables "year", "month",and  "day. The resulting data set **"pols_df** had **`r nrow(pols_df)`** observations, **`r ncol(pols_df)` variables**. The variables were **`r colnames(pols_df)`** and the range of years was **`r range(pols_df$year)`**

* The dataset **snp.csv** describes the closing values of S&P stock index at a given time period. It initially had 787 observations and 2 variables. No columns or observations  were added or removed. The resulting data set **"snp_df** had **`r nrow(snp_df)` observations** and **`r ncol(snp_df)` variables**. The variables were **`r colnames(snp_df)`** and the range of years was **`r range(snp_df$year)`**

* The dataset **unemployment.csv** provides information about percentage of unemployment at a given time. It initially had 68 observations and 13 variables.The dataset was switched from a wide to a long format. The resulting dataset **"unemp_df"** had **`r nrow(unemp_df)` observations** and **`r ncol(unemp_df)` variables**. The variables were **`r colnames(unemp_df)`** and the range of years was **`r range(unemp_df$year)`**

* The **first combined dataset "pomo_snp"** was created using dataset pols_df and snp_df. It had **`r nrow(pomo_snp)` observations** and **`r ncol(pomo_snp)` variables** and the range of years was **`r range(pomo_snp$year, na.rm=TRUE)`** It had variables **`r colnames(pomo_snp)`**. 

* The **second combined dataset "pomo_unemp** was created by merging "pomo_snp" and "unemp_df". It has **`r nrow(pomo_unemp)` observations** and **`r ncol(pomo_unemp)` variables**. The range of years is **`r range(pomo_unemp$year, na.rm=TRUE)`**. It has variables **`r colnames(pomo_unemp)`**

